Glyph-maps for Visualizing Climate Data and Models

### 2011-09-28 paper.tex
spell check, and fig 2 caption

Climate data is challenging to visualise because it is multivariate and has both spatial (2d) and temporal (1d) components. In this paper we propose an new display for climate data: glyph-maps. Glyph-maps are a specialisation of multivariate glyph plots; each spatial location is display with one glyph that represents the measurements recorded over time at that location. Glyph-maps allow the discovery of both local and global structure, making discovery much easier than the classical display using colours.

We include a description of how to create glyph-maps using existing statistical software, our initial explorations of their perceptual features, and a discussion of reference frames. Our techniques are inspired by rectangular gridded data, but we provide generalisation to deal with non-rectangular grids and irregular locations.

@@

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=3.5in]{nasa-deseas-glyph}
  \caption{Glyph-map of de-seasonalized temperature, to compare with Figure \ref{fig:nasa-facet}. }
  \label{fig:nasa-glyph}
\end{figure*}

@@

We can also emphasise different features of the data by varying the scale used with each cell. By default, we use a global scale, so the same position within each cell corresponds to the same value in all locations. This facilitates the comparison of absolute values, and draws attention to plots with large overall variation.

@@

Climate data is usually large: many locations in space, and many points in time. This makes climate data challenging to work with computationally and to visualise. Plotting the raw data often reveals little more than a well-known observation that ``temperature is quite varied!'' Careful use of models than Careful use of models to decompose data into long-term trend, seasonal effects and error is vital for to make the problem both computationally and visually tractable.

>>>>

Climate data is challenging to visualize because it is multivariate and has both spatial (2d) and temporal (1d) components. In this paper we propose an new display for climate data: glyph-maps. Glyph-maps are a specialization of multivariate glyph plots; each spatial location is display with one glyph that represents the measurements recorded over time at that location. Glyph-maps allow the discovery of both local and global structure, making discovery much easier than the classical display using colors.

We include a description of how to create glyph-maps using existing statistical software, our initial explorations of their perceptual features, and a discussion of reference frames. Our techniques are inspired by rectangular gridded data, but we provide generalization to deal with non-rectangular grids and irregular locations.

@@

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=3.5in]{nasa-deseas-glyph}
  \caption{Glyph-map of de-seasonalized temperature, to compare with Figure \ref{fig:nasa-facet}. Time series of the six years of monthly temperature are plotted at each spatial grid location. The El Ni\~no appears as a bump in the middle of the time series, in the equatorial Pacific region. Large variations in temperature can be seen in areas over land, while being fairly constant over water. }
  \label{fig:nasa-glyph}
\end{figure*}

@@

We can also emphasize different features of the data by varying the scale used with each cell. By default, we use a global scale, so the same position within each cell corresponds to the same value in all locations. This facilitates the comparison of absolute values, and draws attention to plots with large overall variation.

@@

Climate data is usually large: many locations in space, and many points in time. This makes climate data challenging to work with computationally and to visualize. Plotting the raw data often reveals little more than a well-known observation that ``temperature is quite varied!'' Careful use of models than Careful use of models to decompose data into long-term trend, seasonal effects and error is vital for to make the problem both computationally and visually tractable.
